### 2020-06-06 report/sections/server/react/efsm.tex
Report: progress on EFSM

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type State = { num1: number, num2: number, message: string }; (*@\label{line:clientsendstatestate}@*)
class InputWindow extends S40<State> {
	render() {
		const { num1, num2, message } = this.state;
		const (*@\hl{Add}@*) = this.ADD('onClick', (*@\label{line:clientsendfactory1}@*)
			(ev: React.MouseEvent) => [num1, num2] (*@\label{line:clientsendpayload1}@*)
		);
		const (*@\hl{Quit}@*) = this.QUIT('onMouseOver', (*@\label{line:clientsendfactory2}@*)
			(ev: React.MouseEvent) => [message] (*@\label{line:clientsendpayload2}@*)
		);		
		return (<div>
			// Omitting <input>s for entering numbers, message
			(*@\hl{<Add>}@*)<button>Sum</button>(*@\hl{</Add>}@*)
			(*@\hl{<Quit>}@*)<button>Quit</button>(*@\hl{</Quit>}@*)
		</div>);
	}
}
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:adderclientsenddev}
\end{figure}

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type FunctionPropertyNames<T> = {
	[K in keyof T]: T[K] extends Function | undefined ? K : never; (*@\label{line:extractfuncprops1}@*)
}[keyof T]; (*@\label{line:extractfuncprops2}@*)
type FunctionProperties<T> = Pick<T, FunctionPropertyNames<T>>; (*@\label{line:extractfuncprops3}@*)
type DOMEvents = FunctionProperties<React.DOMAttributes<any>>;
\end{lstlisting}
\captionof{lstlisting}{\dots}
\label{lst:extractfuncprops}
\end{figure}

@@

\subsection{Receive States}
\begin{itemize}
\item receive state is `shown' when waiting for a receive event
\item base class has abstract methods -- one for each possible receive -- so the user must implement these when they extend the abstract base class
\item each abstract method is named after the message label expected to be received, and parameterised by the payload
\item no channel actions are passed to the state -- instead, once the receive state is mounted on the DOM (i.e. rendered), the abstract base class will register the event handlers to the runtime (using the register method passed as props)
\item when the message arrives, the runtime will invoke the registered function (+ forward reference the next section with the edge case that the message arrives before handler registration occurs)
\end{itemize}

@@

>>>>

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type State = { num1: number, num2: number, message: string }; (*@\label{line:clientsendstatestate}@*)
class InputWindow extends S40<State> {
	render() {
		const { num1, num2, message } = this.state;
		const (*@\hl{Add}@*) = this.ADD('onClick', (*@\label{line:clientsendfactory1}@*)
			(ev: React.MouseEvent) => [num1, num2]); (*@\label{line:clientsendpayload1}@*)
		const (*@\hl{Quit}@*) = this.QUIT('onMouseOver', (*@\label{line:clientsendfactory2}@*)
			(ev: React.MouseEvent) => [message]); (*@\label{line:clientsendpayload2}@*)
		return (<div>
			// Omitting <input>s for entering numbers, message
			(*@\hl{<Add>}@*)<button>Sum</button>(*@\hl{</Add>}@*)
			(*@\hl{<Quit>}@*)<button>Quit</button>(*@\hl{</Quit>}@*)
		</div>);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Developer Implementation for \trole{Client}
Send State in \tprotocol{Adder} protocol}
\label{lst:adderclientsenddev}
\end{figure}

@@

\begin{figure}[!h]
\begin{lstlisting}[language=javascript,tabsize=2]
type FunctionPropertyNames<T> = {
	[K in keyof T]: T[K] extends Function | undefined ? K : never; (*@\label{line:extractfuncprops1}@*)
}[keyof T]; (*@\label{line:extractfuncprops2}@*)
type FunctionProperties<T> = Pick<T, FunctionPropertyNames<T>>; (*@\label{line:extractfuncprops3}@*)
type DOMEvents = FunctionProperties<React.DOMAttributes<any>>;
\end{lstlisting}
\captionof{lstlisting}{Extracting Function Properties from 
TypeScript Interface}
\label{lst:extractfuncprops}
\end{figure}

@@

\subsection{Receive States}

We also encode receive states as abstract React components,
and leverage props to ensure channel resources are not exposed
to the developer.

According to the EFSM, a receive state means
the endpoint is ``waiting for the message'',
and the actual receive event represents a transition
to a successor state.
This means that the developer's implementation will define
what to render on the DOM when the endpoint is waiting
for the message, rather than updating the DOM with the
received message.
Referring to \cref{fig:adderclientfsm} again,
S42 will be rendered when the endpoint is waiting for
\tmsg{RES}, but once received, S40 will be immediately
rendered. The developer will need to manage the flow
of stateful application data between UI components separately
-- \fancyname{ReactMPST} only generates code for managing
the communication automaton.

We generate an abstract method for each permitted receive
transition, which is used as the receive handler.
This allows the developer to intercept the
message and perform operations -- such as updating
application state -- before the successor state is rendered.
Specifying receive handlers as abstract methods mandate
the developer to provide a concrete implementation.

\begin{lstlisting}[language=javascript,tabsize=2]
enum Labels {
	THANKS = 'THANKS', TERMINATE = 'TERMINATE',
};

type THANKSMessage = {
    label: Labels.THANKS,
    payload: [string],
};
type TERMINATEMessage = {
    label: Labels.TERMINATE,
    payload: [],
};

type Message = | THANKSMessage | TERMINATEMessage;
\end{lstlisting}

\begin{lstlisting}[language=javascript,tabsize=2]
type Props = {
	register: (role: Roles.Peers, handle: ReceiveHandler) => void
};
abstract class S42<S> extends React.Component<Props, S> {
	abstract THANKS(payload1: string): void;
	abstract TERMINATE(): void;

	componentDidMount() {
		this.props.register(this.handle.bind(this));	
	}

	handle(message: any): State {
		const parsedMessage = JSON.parse(message) as Message;
		switch (parsedMessage.label) {
			case Labels.THANKS:
				return this.THANKS(...parsedMessage.payload);			
			case Labels.TERMINATE:
				return this.TERMINATE(...parsedMessage.payload);
		}
	}
}
\end{lstlisting}

\begin{itemize}
\item receive state is `shown' when waiting for a receive event
\item base class has abstract methods -- one for each possible receive -- so the user must implement these when they extend the abstract base class
\item each abstract method is named after the message label expected to be received, and parameterised by the payload
\item no channel actions are passed to the state -- instead, once the receive state is mounted on the DOM (i.e. rendered), the abstract base class will register the event handlers to the runtime (using the register method passed as props)
\item when the message arrives, the runtime will invoke the registered function (+ forward reference the next section with the edge case that the message arrives before handler registration occurs)
\end{itemize}

@@

\begin{lstlisting}[language=javascript,tabsize=2]
type Props = { terminate: () => void };
abstract class S41<S> extends React.Component<Props, S> {
	componentDidMount() { this.props.terminate(); }	
}
\end{lstlisting}
